---
title: "Coffee Quality"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidytuesdayR)
library(maps)
library(ggplot2)
library(tidyr)
library(dplyr)
library(plotly)
library(patchwork)
```




```{r global, include=FALSE}
# function to test if the directory exists, and create one if not
data_dir <- function(folder) {
  if(!file.exists(paste(here::here(), folder, sep="/"))) {
    dir.create(paste(here::here(), folder, sep="/"))
    }
}
data_dir("data")

# write.csv(coffee_ratings, 
#           file=paste(here::here(), 'data/coffee_ratings.csv', sep="/"), 
#           row.names = FALSE)

data <- read.csv(paste(here::here(), 'data/coffee_ratings.csv', sep="/"))


world <- map_data("world")

world$country_of_origin <- world$region
world[world$region == "USA" & !is.na(world$subregion) & world$subregion == "Hawaii",
      "country_of_origin"] <- "United States (Hawaii)"
world[world$region == "Puerto Rico","country_of_origin"] <- "United States (Puerto Rico)"
world[world$region == "USA" & !is.na(world$subregion) & world$subregion != "Hawaii","country_of_origin"] <- "United States"
world[world$region == "USA" & is.na(world$subregion),"country_of_origin"] <- "United States"
world[world$region == "Tanzania","country_of_origin"] <- "Tanzania, United Republic Of"


# by code
data.code <- data

iso <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv")[, c("COUNTRY", "CODE")]
iso <- rbind(iso, c("Myanmar", "MMR"))

data.code[!is.na(data.code$country_of_origin) & data.code$country_of_origin == "Cote d?Ivoire","country_of_origin"] <- "Cote d'Ivoire"
data.code[!is.na(data.code$country_of_origin) & data.code$country_of_origin == "Tanzania, United Republic Of","country_of_origin"] <- "Tanzania"
data.code[!is.na(data.code$country_of_origin) & data.code$country_of_origin == "United States (Hawaii)","country_of_origin"] <- "United States"
data.code[!is.na(data.code$country_of_origin) & data.code$country_of_origin == "United States (Puerto Rico)","country_of_origin"] <- "Puerto Rico"
```

# About

The data is on coffee <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md>

Coffee Quality Database

```{r, eval = FALSE}
tuesdata <- tidytuesdayR::tt_load('2020-07-07')
tuesdata <- tidytuesdayR::tt_load(2020, week = 28)

coffee_ratings <- tuesdata$coffee_ratings
```


# Data

```{r}
data %>% 
  head(100) %>% 
  select(total_cup_points, species, owner, country_of_origin, company,
         processing_method, aroma, flavor, aftertaste, acidity, altitude_mean_meters) %>% 
  DT::datatable()
```




# Static 1 {data-navmenu="Static"}

## Column {data-width=550}

### test LOL

```{r}
data %>% 
  filter(total_cup_points != 0) %>%
  ggplot(aes(x=total_cup_points)) + geom_histogram() + 
  ggtitle("Histogram of Total Cup Points", subtitle = "Mean and Median included") + 
  geom_vline(aes(xintercept=mean(total_cup_points)),
            color="blue", linetype="dashed", size=1) + 
  geom_vline(aes(xintercept=median(total_cup_points)),
            color="red", linetype="dashed", size=1)
```

## Column {data-width=450}

### Chart B

```{r}
p1 <- ggplot(data, aes(x=factor(0), y=altitude_mean_meters)) + 
  geom_boxplot() + ggtitle("Boxplot of mean altitude (all coffee included)") + ylab("mean altitude (meters)") + xlab("") + coord_flip()
p2 <- data %>%
  filter(altitude_mean_meters < 10000) %>%
  ggplot(aes(x=factor(0), y=altitude_mean_meters)) + 
  geom_boxplot() + ggtitle("Boxplot of mean altitude with abnormal removed") + ylab("mean altitude (meters)")  + xlab("") + coord_flip()
p1/p2 + plot_annotation(
  title = "Boxplot of mean altitude in meters"
)
```

### Chart C

```{r}
data %>% 
  filter(total_cup_points != 0) %>%
  filter(altitude_mean_meters < 10000) %>%
  ggplot(aes(x=altitude_mean_meters, y=total_cup_points)) + 
  geom_point() + geom_smooth()  + ggtitle("Cup points against Altitude", subtitle = "Trendline included") + ylab("Total Cup Points")
```

# Static 2 {data-navmenu="Static"}

## Test {.tabset}

### Body

```{r}
df <- data.code %>%
  group_by(country_of_origin) %>%
  summarise(mean_variable = mean(body, na.rm = TRUE)) %>%
  full_join(iso, by = c("country_of_origin" = "COUNTRY"))

g <- list(
  showframe = FALSE,
  showcoastlines = TRUE,
  lonaxis = list(showgrid = T),
  lataxis = list(showgrid = T),
  projection = list(type = 'Mercator')
)

fig <- plot_geo(df)
l <- list(color = toRGB("grey"), width = 0.5)
fig <- fig %>% add_trace(
    z = ~mean_variable, color = ~mean_variable, colors = 'Blues',
    text = ~country_of_origin, locations = ~CODE, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'Body')
fig <- fig %>% layout(
    title = 'Geographic Distribution of Mean Coffee Body',
    geo = g
  )

fig
```

### Sweetness

```{r}
df <- data.code %>%
  group_by(country_of_origin) %>%
  summarise(mean_variable = mean(sweetness, na.rm = TRUE)) %>%
  full_join(iso, by = c("country_of_origin" = "COUNTRY"))

g <- list(
  showframe = FALSE,
  showcoastlines = TRUE,
  lonaxis = list(showgrid = T),
  lataxis = list(showgrid = T),
  projection = list(type = 'Mercator')
)

fig <- plot_geo(df)
l <- list(color = toRGB("grey"), width = 0.5)
fig <- fig %>% add_trace(
    z = ~mean_variable, color = ~mean_variable, colors = 'Blues',
    text = ~country_of_origin, locations = ~CODE, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'Sweetness')
fig <- fig %>% layout(
    title = 'Geographic Distribution of Mean Coffee Sweetness',
    geo = g
  )

fig
```


### Acidity

```{r}
df <- data.code %>%
  group_by(country_of_origin) %>%
  summarise(mean_variable = mean(acidity, na.rm = TRUE)) %>%
  full_join(iso, by = c("country_of_origin" = "COUNTRY"))

g <- list(
  showframe = FALSE,
  showcoastlines = TRUE,
  lonaxis = list(showgrid = T),
  lataxis = list(showgrid = T),
  projection = list(type = 'Mercator')
)

fig <- plot_geo(df)
l <- list(color = toRGB("grey"), width = 0.5)
fig <- fig %>% add_trace(
    z = ~mean_variable, color = ~mean_variable, colors = 'Blues',
    text = ~country_of_origin, locations = ~CODE, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'Acidity')
fig <- fig %>% layout(
    title = 'Geographic Distribution of Mean Coffee Acidity',
    geo = g
  )

fig
```


### Flavor

```{r}
df <- data.code %>%
  group_by(country_of_origin) %>%
  summarise(mean_variable = mean(flavor, na.rm = TRUE)) %>%
  full_join(iso, by = c("country_of_origin" = "COUNTRY"))

g <- list(
  showframe = FALSE,
  showcoastlines = TRUE,
  lonaxis = list(showgrid = T),
  lataxis = list(showgrid = T),
  projection = list(type = 'Mercator')
)

fig <- plot_geo(df)
l <- list(color = toRGB("grey"), width = 0.5)
fig <- fig %>% add_trace(
    z = ~mean_variable, color = ~mean_variable, colors = 'Blues',
    text = ~country_of_origin, locations = ~CODE, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'Flavor')
fig <- fig %>% layout(
    title = 'Geographic Distribution of Mean Coffee Flavor',
    geo = g
  )

fig
```


# Interactive 1 {data-navmenu="Interactive"}

## Column {.sidebar}

```{r}
selectInput("cup", "Cupping Criteria:",
                  c("Flavor" = "flavor",
                    "Sweetness" = "sweetness",
                    "Body" = "body",
                    "Acidity" = "acidity",
                    "Aroma" = "aroma",
                    "Balance" = "balance",
                    "Uniformity" = "uniformity",
                    "Clean Cup" = "clean_cup",
                    "Moisture" = "moisture"))
```


## Column

```{r}
formulaText1 <- reactive({
    paste("total_cup_points ~ ", input$cup)
  })

formulaTitle1 <- reactive({
    paste("Total Cup Points versus", 
          gsub("_", " " , input$variable2), 
          "with correlation", 
          round(cor(data$total_cup_points, data[,input$cup]),2)
          )
  })

renderPlot({
    plot(as.formula(formulaText1()),
         data = data[data$total_cup_points != 0,],
         pch = 20,
         col = "cyan4",
         main = formulaTitle1(),
         ylab = "Total Cup Points",
         xlab = gsub("_", " " , input$variable2))
  abline(lm(as.formula(formulaText1()), data = data[data$total_cup_points != 0,]), 
         col = "coral1", lwd = 2)
  })
```

# Interactive 2 {data-navmenu="Interactive"}


## Column {.sidebar}

acidity, aroma, flavor, aftertaste, body, balance, uniformity, clean_cup, sweetness, moisture


```{r}
selectInput("variable1", "Cupping Criteria:",
                  c("Flavor" = "flavor",
                    "Sweetness" = "sweetness",
                    "Body" = "body",
                    "Acidity" = "acidity",
                    "Aroma" = "aroma",
                    "Balance" = "balance",
                    "Uniformity" = "uniformity",
                    "Clean Cup" = "clean_cup",
                    "Moisture" = "moisture"))

selectInput("variable2", "Methods:",
                  c("Processing Method" = "processing_method",
                    "Species" = "species"))

checkboxInput("outliers", "Show outliers", TRUE)
```

## Column

### bar

```{r}
formulaText <- reactive({
    paste(input$variable1, " ~", input$variable2)
  })

formulaTitle <- reactive({
    paste(gsub("_", " " , input$variable1), "versus", gsub("_", " " , input$variable2))
  })

  # Generate a plot of the requested variable against mpg ----
  # and only exclude outliers if requested
renderPlot({
    boxplot(as.formula(formulaText()),
            data = data,
            outline = input$outliers,
            col = "#75AADB", pch = 19, main=formulaTitle())
  })
```


```{r, eval=FALSE, include=FALSE}
renderPlot({
data %>%
  filter(!is.na(!!sym(input$variable))) %>%
  ggplot(aes_string(x=input$variable2, y=input$variable)) +
geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 45,  hjust=1)) + ggtitle("test")
})
```

# Analysis

## Tables {.tabset} {data-width=450}

### Min cup points

```{r}
data %>% 
  slice_min(total_cup_points, n=5) %>%
  select(total_cup_points, species, acidity, country_of_origin, processing_method, altitude_mean_meters) %>%
  DT::datatable()
```
### top cup points

```{r}
data %>% 
  slice_max(total_cup_points, n=5) %>%
  select(total_cup_points, species, acidity, country_of_origin, processing_method, altitude_mean_meters) %>%
  arrange(desc(total_cup_points)) %>%
  DT::datatable()
```

### Summary of cup point attributes by processing method

```{r}
data %>%
  filter(!is.na(total_cup_points)) %>%
  filter(!is.na(processing_method)) %>%
  group_by(processing_method) %>%
  summarise(mean_cup_points = mean(total_cup_points, na.rm = T),
            mean_acidity = mean(acidity, na.rm = T),
            mean_aroma = mean(aroma, na.rm = T),
            mean_flavor = mean(flavor, na.rm = T),
            mean_aftertaste = mean(aftertaste, na.rm = T),
            mean_body = mean(body, na.rm = T),
            mean_balance = mean(balance, na.rm = T),
            mean_uniformity = mean(uniformity, na.rm = T),
            mean_clean_cup = mean(clean_cup, na.rm = T),
            mean_sweetness = mean(sweetness, na.rm = T),
            mean_moisture = mean(moisture, na.rm = T)) %>%
  DT::datatable()
```

### Summary of cup point attributes by coffee species


```{r}
data %>%
  filter(!is.na(species)) %>%
  group_by(species) %>%
  summarise(mean_cup_points = mean(total_cup_points, na.rm = T),
            mean_acidity = mean(acidity, na.rm = T),
            mean_aroma = mean(aroma, na.rm = T),
            mean_flavor = mean(flavor, na.rm = T),
            mean_aftertaste = mean(aftertaste, na.rm = T),
            mean_body = mean(body, na.rm = T),
            mean_balance = mean(balance, na.rm = T),
            mean_uniformity = mean(uniformity, na.rm = T),
            mean_clean_cup = mean(clean_cup, na.rm = T),
            mean_sweetness = mean(sweetness, na.rm = T),
            mean_moisture = mean(moisture, na.rm = T)) %>%
  DT::datatable()
```

## Plots {.tabset} {data-width=550}

### altitude

```{r}
p1 <- ggplot(data, aes(x=factor(0), y=altitude_mean_meters)) + 
  geom_boxplot() + ggtitle("Boxplot of mean altitude (all coffee included)") + ylab("mean altitude (meters)") + xlab("") + coord_flip()
p2 <- data %>%
  filter(altitude_mean_meters < 10000) %>%
  ggplot(aes(x=factor(0), y=altitude_mean_meters)) + 
  geom_boxplot() + ggtitle("Boxplot of mean altitude with abnormal removed") + ylab("mean altitude (meters)")  + xlab("") + coord_flip()
p1/p2 + plot_annotation(
  title = "Boxplot of mean altitude in meters"
)
```

### altitude vs cup points

```{r}
data %>% 
  filter(total_cup_points != 0) %>%
  filter(altitude_mean_meters < 10000) %>%
  ggplot(aes(x=altitude_mean_meters, y=total_cup_points)) + 
  geom_point() + geom_smooth()  + ggtitle("Cup points against Altitude", subtitle = "Trendline included") + ylab("Total Cup Points")
```

### batplot of acidity


```{r}
data %>%
  filter(!is.na(processing_method)) %>%
  ggplot(aes(x=processing_method, y=acidity)) +
geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 45,  hjust=1)) + ggtitle("Barplot of Acidity versus processing method")
```

### After Taste

```{r}
data %>%
  filter(data$total_cup_points != 0) %>%
  ggplot(aes(x=aftertaste, y=total_cup_points)) + 
  geom_point()  + facet_grid(cols = vars(species))  + ggtitle("Total Cup Points versus After Taste", subtitle = "grouped by species") + xlab("After Taste") + ylab("Total Cup Points")
```


### Data with code

```{r}
df <- data.code %>%
  group_by(country_of_origin) %>%
  summarise(mean_acidity = mean(acidity, na.rm = TRUE)) %>%
  full_join(iso, by = c("country_of_origin" = "COUNTRY"))

g <- list(
  showframe = TRUE,
  showcoastlines = TRUE,
  projection = list(type = 'Mercator')
)

fig <- plot_geo(df)
l <- list(color = toRGB("grey"), width = 0.5)
fig <- fig %>% add_trace(
    z = ~mean_acidity, color = ~mean_acidity, colors = 'Blues',
    text = ~country_of_origin, locations = ~CODE, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'Mean Acidity')
fig <- fig %>% layout(
    title = 'Mean Acidity by location',
    geo = g
  )

fig
```
  
```{r}
data.code %>%
  group_by(country_of_origin) %>%
  summarise(mean_acidity = mean(acidity, na.rm = TRUE)) %>%
  full_join(iso, by = c("country_of_origin" = "COUNTRY")) %>%
  plot_ly(type='choropleth', locations=~CODE, z=~mean_acidity, text=~country_of_origin, colorscale="Blues")
```

### data with coordinate

```{r}
data %>% 
  group_by(country_of_origin) %>%
  summarise(mean_balance = mean(balance, na.rm = TRUE)) %>%
  full_join(world, by = "country_of_origin") %>%
  ggplot(aes(long, lat)) +
  geom_polygon(aes(group = group, fill = mean_balance)) +
  scale_fill_continuous(low="thistle2", high="darkred", 
                       guide="colorbar",na.value="white") +
  ggtitle("Geographical Distribution of Mean Coffee Balance") + 
  labs(fill = "Mean Acidity")
```

### Functions Used

| Package    | Functions |
| -- | ----- |
| `dplyr` | `slice_max()`, `filter()`, `arrange()`, `full_join()`, `slice_min()`, `summarise()`, `group_by()`|
| `ggplot2` |  `geom_point()`, `geom_histogram()`, `geom_vline()`, `geom_smooth()`, `geom_boxplot()`, `geom_polygon()` |


